A New Method of Modelling Tuneable Lasers with Functional Composition

### 2019-05-22 Thesis/Intro/Introduction.tex
Made some edits.

The word laser was originally an acronym for \emph{Light Amplification by Stimulated Emission of Radiation}. The cavity of a standard laser---such as a HeNe laser---is composed of four main components; a highly reflecting mirror, a partially reflecting mirror, something to supply the pumping energy, and the gain medium. The two mirrors are at opposite ends of the cavity and serve to contain most of the light to the cavity while allowing a small amount to exit, this is the output of the laser. The pump energy is generally supplied by either an external light source, or electric field. The gain medium is what amplifies the laser light and consists of gases, such as Helium and Neon, or as a crystal, such as Neodymium-doped Yttrium Aluminium garnet (Nd:YAG). \\

More precisely, the atoms in the gain medium are pumped with energy raising the electrons to an excited state. When a photon in the laser cavity interacts with such an electron it becomes amplified by stimulated emission, these two process are shown in Figure \ref{fig:emission}. The amplified light perpetually bounces between the two mirrors becoming more amplified with each interaction with an excited electron, a fraction of this light is able to pass through the partially reflecting mirror to become the output of the laser. \\

This is how laser light is generated, but, laser light has two fundamental characteristics that regular light does not. The first is that the light is highly monochromatic---the light contains (ideally) a single frequency. This is in contrast with an incandescent light bulb, for example, which emits light at all wavelengths with relative intensities given by Planck's law of black-body radiation. The other key feature is coherence, in which all the peaks and troughs of the light overlap giving very strong constructive interference---this is why laser light can be so intense. Something like a light emitting diode (LED) is not laser light because, while monochromatic, the light is not coherent. The monochromatic nature is a result of how the light is generated, the frequency of the light emitted directly corresponds to the energy difference between the excited and ground states of the gain medium (as with LEDs). Note that this correspondence means that  the frequency the laser operates at is fixed, and cannot be changed. Coherence is achieved in the laser cavity---the cavity is carefully designed so the light resonates with the length of the cavity. This causes all the peaks and troughs to align, and interfere constructively. \\

\section{Tuneable Lasers}
\begin{figure}[tbp]
\centering
\input{./Figures/Cavity}
\caption{Typical cavity of a tuneable laser.}
\label{fig:cavity}
\end{figure}

A tuneable laser on the other hand, has the ability to vary its wavelength by about 100 nanometres \cite{bohun, burgoyne2010, yamashita}. Tuneable lasers in fact lase at all of these different frequencies simultaneously. This tuneability is quite useful and has applications in spectroscopy and high resolution imaging such as coherent anti-Stokes Raman spectroscopy and optical coherence tomography \cite{bohun, burgoyne2014, yamashita}. However, having the capacity to tune the wavelength comes with a cost, tuneable lasers vary considerably from standard lasers in both construction and operation, but, the central ideas are the same. A typical tuneable laser cavity is shown in Figure \ref{fig:cavity}. Instead of being perpetually reflected back and forth, the light travels in a circular fashion around the cavity, this is known as a ring laser. The gain medium is still excited by an external power source, and the output passes through a partially reflecting mirror to become the output. However, a tuneable laser also contains a few additional components, all of which will be described in the following subsections.

@@

Moreover, a Mach--Zehnder interferometer can be used in tandem with an acousto- or electro-optic modulator to modulate an optical pulse. The incident pulse is separated with a Y-waveguide, one of the two branches is then modulated, finally the two branches are recombined with another Y-waveguide \cite{alazzawi, hausbook, karim}. In the absence of a modulator on one of the branches the two pulses recombine into the original pulse. However, with modulation the two pulses are no longer identical, and so when the recombine, they interfere. This interference can be used to further modulate the pulse \cite{agrawal2002, agrawal2013, hausbook, karim}. \\

@@

\begin{figure}[tbp]
\centering
\input{./Figures/CFBG}
\caption[CFBG]{CFBG}
\label{fig:cfbg}
\end{figure}

The speed of light in an optical fibre is slightly dependent on the wavelength---this causes light with a longer wavelength to travel faster, and is known as chromatic dispersion. This is a large problem in fibre optic communications, the signal can spread and potentially becomes uninterpretable after large distances. However, chromatic dispersion can be counteracted using a CFBG \cite{agrawal2002, alazzawi, becker, starodoumov} (in the opposite orientation of Figure \ref{fig:cfbg}). By forcing the longer wavelengths to travel farther the dispersion can be reversed, restoring the original signal. In one experiment \cite{dong}, a signal was successfully transmitted over $109$ km at $40$ Gb/s by compensating the dispersion with two  $40$ cm CFBGs. Over this distance, the pulse would have spread to about $55$ times its original width, and could only have been transmitted $4$ km at that bit rate. At reduced bit rates however, a $10$ cm CFBG can compensate the dispersion of $300$ km of fibre \cite{agrawal2002}. However, in our case, we wish to accelerate the dispersion, simulating hundreds of metres of fibre, so the CFBG is used in the orientation shown in Figure \ref{fig:cfbg}. \\

@@

\subsection{Optical Amplifier and Pump Laser}
Optical amplifiers are of particular importance in fibre optic communications, they are used to restore the strength of a signal after it hass been attenuated over large distances, or when a signal is divided into multiple paths \cite{alazzawi, starodoumov}. They are also more efficient, and introduce less noise than an electrical repeater. In this context, the optical amplifier provides the energy of the laser \cite{alazzawi}. Most commonly, optical amplifiers are created  doping a length of fibre (called the gain fibre) with a rare-earth element which receives power from a pump laser \cite{agrawal2002, alazzawi, starodoumov}. The most common dopant is Erbium, however, Ytterbium and Neodymium are also used---Holmium, Samarium, Thulium, and Tellurium are infrequently used as well \cite{agrawal2002}. The amplification is achieved with stimulated emission---a similar process to spontaneous emission. \\

Spontaneous emission is the process in which an electron of an atom that is in an excited state emits a photon as it transitions to a lower energy state. The energy of the photon is equal to the energy difference between the two levels. On the other hand, in stimulated emission an incident photon triggers this emission \cite{alazzawi}. The rare-earth metal ions are excited by the pump laser, and when they interact with a photon of the tuneable laser, a new photon is released with the same direction and phase as the incident one, maintaining coherence \cite{alazzawi}. These processes are highlighted in Figure \ref{fig:emission}. \\

@@

\subsection{Mode-Locking}
\begin{figure}[tbp]
\centering
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML1}
\caption{$n = 4$}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML2}
\caption{$n = 12$}
\end{subfigure}
\caption{Two examples of mode-locking given by $\displaystyle \sum_{i = 1}^n \cos \left( 2 \pi i x \right)$.}
\label{fig:ml}
\end{figure}
One of the fundamental features of a laser is coherence. However, in the context of tuneable lasers it is perhaps less clear what is meant by coherence since the light is no longer monochromatic. Instead we have a particular peak of all the frequencies aligned, this is called mode-locking---all of the different modes are locked together at a peak. This of course creates a very intense and short pulse of light, at this particular peak all of the frequencies constructively interfere as with a standard laser, however, deviations from the peak lead to destructive interference. Two examples of this are shown in Figure \ref{fig:ml}, and as expected, including more modes narrows and intensifies the pulse. \\

how it's done

>>>>

The word laser was originally an acronym for \emph{Light Amplification by Stimulated Emission of Radiation}. The cavity of a standard laser---such as a HeNe laser---is composed of four main components: a highly reflective mirror, a partially reflecting mirror, something to supply the pumping energy, and the gain medium. The two mirrors are at opposite ends of the cavity and serve to contain most of the light to the cavity while allowing a small amount to exit, this is the output of the laser. The pump energy is generally supplied by either an external light source, or an electric field. The gain medium is what amplifies the laser light and consists of gases, such as Helium and Neon, or a crystal, such as Neodymium-doped Yttrium Aluminium garnet (Nd:YAG). \\

More precisely, the atoms in the gain medium are pumped with energy raising the electrons to an excited state. For stimulated emission an incident photon triggers the emission of a photon from this excited electron \cite{alazzawi}. This new photon absorbs the electron's energy as it returns to its ground state, amplifying the light. Furthermore, the photon is released with the same direction and phase as the incident one, maintaining coherence \cite{alazzawi}. These processes are highlighted in Figure \ref{fig:emission}. The amplified light perpetually bounces between the two mirrors becoming more amplified with each interaction with an excited electron, a fraction of this light is able to pass through the partially reflecting mirror to become the output of the laser. \\

This is how laser light is generated, but, laser light has two fundamental characteristics that regular light does not. The first is that the light is highly monochromatic---the light contains (ideally) a single frequency. This is in contrast with an incandescent light bulb, for example, which emits light at all wavelengths with relative intensities given by Planck's law of black-body radiation. The other key feature is coherence, in which all the peaks and troughs of the light overlap giving very strong constructive interference---this is why laser light can be so intense. Something like a light emitting diode (LED) does not produce laser light because, while monochromatic, the light is not coherent. The monochromatic nature is a result of how the light is generated, the frequency of the light emitted directly corresponds to the energy difference between the excited and ground states of the gain medium (as with LEDs). Note that this correspondence means that  the frequency the laser operates at is fixed, and cannot be changed. Coherence is achieved in the laser cavity---the cavity is carefully designed so the light resonates with the length of the cavity. This causes all the peaks and troughs to align, and interfere constructively. \\

\section{Tuneable Lasers}
\begin{figure}[tbp]
\centering
\input{./Figures/Cavity}
\caption[Typical cavity of a tuneable laser.]{Typical cavity of a tuneable laser. The pulses travel clockwise around the loop.}
\label{fig:cavity}
\end{figure}

A tuneable laser on the other hand, has the ability to vary its wavelength by up to about 100 nanometres \cite{bohun, burgoyne2010, yamashita}. Tuneable lasers in fact lase at all of these different frequencies simultaneously. This tuneability is quite useful and has applications in spectroscopy and high resolution imaging such as coherent anti-Stokes Raman spectroscopy and optical coherence tomography \cite{bohun, burgoyne2014, yamashita}, as well as communications and diagnostics of ultra fast processes \cite{silfvast}. However, having the capacity to tune the wavelength comes with a cost, tuneable lasers vary considerably from standard lasers in both construction and operation, but, the central ideas are the same. A typical tuneable laser cavity is shown in Figure \ref{fig:cavity}. Instead of being perpetually reflected back and forth, the light travels in a circular fashion around the cavity, this is a so called ring laser. The gain medium is still excited by an external power source, and the output passes through a partially reflecting mirror to become the output. However, a tuneable laser also contains a few additional components, each of which will be described in the following subsections.

@@

Moreover, a Mach--Zehnder interferometer can be used in tandem with an acousto- or electro-optic modulator to modulate an optical pulse. The incident pulse is separated with a Y-waveguide, one of the two branches is then modulated, and finally the two branches are recombined with another Y-waveguide \cite{alazzawi, hausbook, karim}. In the absence of a modulator on one of the branches the two pulses recombine into the original pulse. However, with modulation the two pulses are no longer identical, and so when the recombine, they interfere. This interference can be used to further modulate the pulse \cite{agrawal2002, agrawal2013, hausbook, karim}. \\

@@

\begin{figure}[tbp]
\centering
\input{./Figures/CFBG}
\caption[Chirped fibre Bragg grating.]{Chirped fibre Bragg grating. Light is reflected when the wavelength coincides with the Bragg wavelength. This causes light with different wavelengths to penetrate to different depths. In turn, dispersion is either heightened or compensated---depending on the orientation.}
\label{fig:cfbg}
\end{figure}

Furthermore, the speed of light in an optical fibre is slightly dependent on the wavelength---this causes light with a longer wavelength to travel faster, and is known as chromatic dispersion. This is a large problem in fibre optic communications, the signal can spread and potentially becomes uninterpretable after large distances. However, chromatic dispersion can be counteracted using a CFBG \cite{agrawal2002, alazzawi, becker, starodoumov} (in the opposite orientation of Figure \ref{fig:cfbg}). By forcing the longer wavelengths to travel farther the dispersion can be reversed, restoring the original signal. In one experiment \cite{dong}, a signal was successfully transmitted over $109$ km at $40$ Gb/s by compensating the dispersion with two  $40$ cm CFBGs. Over this distance, the pulse would have spread to about $55$ times its original width, and could only have been transmitted $4$ km at that bit rate. At reduced bit rates however, a $10$ cm CFBG can compensate the dispersion of $300$ km of fibre \cite{agrawal2002}. However, in our case, we wish to accelerate the dispersion, simulating hundreds of metres of fibre, so the CFBG is used in the orientation shown in Figure \ref{fig:cfbg}. \\

@@

\subsection{Optical Amplifier and Pump Laser}
Optical amplifiers are of particular importance in fibre optic communications, they are used to restore the strength of a signal after it has been attenuated over large distances, or when a signal is divided into multiple paths \cite{alazzawi, starodoumov}. They are also more efficient, and introduce less noise than an electrical repeater. In this context, the optical amplifier provides the energy of the laser \cite{alazzawi}. Most commonly, optical amplifiers are created by doping a length of fibre (called the gain fibre) with a rare-earth element which receives power from a pump laser \cite{agrawal2002, alazzawi, starodoumov}. The most common dopant is Erbium, however, Ytterbium and Neodymium are also used---Holmium, Samarium, Thulium, and Tellurium are infrequently used as well \cite{agrawal2002}. \\

@@

\subsection{Mode-Locking}
\begin{figure}[tbp]
\centering
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML1}
\caption{$n = 4$}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML2}
\caption{$n = 12$}
\end{subfigure}
\caption[Two examples of mode-locking.]{Two examples of mode-locking given by $\displaystyle \sum_{i = 1}^n \cos \left( 2 \pi i x \right)$.}
\label{fig:ml}
\end{figure}
One of the fundamental features of a laser is coherence. However, in the context of tuneable lasers it is perhaps less clear what is meant by coherence since the light is no longer monochromatic. Instead, we have a particular peak of all the frequencies aligned, that is, the phase shift between modes is zero. This process is called mode-locking---all of the different modes are locked together at a peak \cite{agrawal2013, hausbook, silfvast}. This of course creates a very intense and short pulse of light---as short as $5$ fs \cite{silfvast}. At this particular peak all of the frequencies constructively interfere as with a standard laser, however, deviations from the peak lead to destructive interference. Two examples of this are shown in Figure \ref{fig:ml}, and as expected, including more modes narrows and intensifies the pulse. \\

### 2019-05-22 Thesis/Model/Model.tex
Made some edits.

\chapter{A New Model}
\label{chap:linear}
In this chapter we shall derive our new model and solve it analytically in the linear case. To accomplish this we shall use the ideas of the previous functional models \cite{cutler, siegman, kuizenga1970a, kuizenga1970b, kuizenga1970, martinez1984, martinez1985, burgoyne2014}. To alleviate some of the pitfalls mentioned in the previous chapter, we shall include all five processes involved in the modification of the pulse within the laser cavity (gain, nonlinearity, loss, dispersion, and modulation). In addition to this, the functional operations associated with each component will be derived from \eqref{eq:nlse}, with the exception of the modulation in which we consider the exact functional form to be determined by the laser operator. \\

We shall start our analysis with the derivation of the functional operators for the five components. \\

\section{Components}
\subsection{Gain}
\label{chap:gain}
Within the Er-doped gain fibre, the gain term is dominant, and equation~\eqref{eq:nlse} reduces to
\begin{align}
\label{eq:gainde}
\pdiff{A}{z} &= \frac{1}{2} g(A),
\end{align}
where $g(A)$ takes the form \cite{bohun, burgoyne2014, hausbook, haus1975, haus1991, haus1992, haus2000, peng, shtyrina, silfvast, usechak, yarutkina}
\begin{align}
\label{eq:energy}
	g(A) &= \frac{g_0}{1 + E / \Es} A,& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate.
Multiplying~\eqref{eq:gainde} by $\bar{A}$, the complex conjugate of $A$, yields
\begin{align*}
	\bar{A} \pdiff{A}{z} = \frac{1}{2}\frac{g_0 |A|^2}{1 + E / \Es},
\end{align*}
adding this to its complex conjugate and integrating over $T$ gives
%\begin{align*}
%	\diff{|A|^2}{z} &= \frac{g_0 |A|^2}{1 + E / \Es}.
%\end{align*}
%After integrating this becomes
\begin{align}
\label{diffez}
	\diff{E}{z} &= \frac{g_0 E}{1 + E / \Es}.
\end{align}
For $E \ll \Es$ the energy grows exponentially, whereas for $E \gg \Es$ the gain has saturated and so the growth is linear. To obtain a closed form solution, \eqref{diffez} is integrated over a gain fibre of length $z$ and assuming the energy increases from $E$ to $E_{\textrm{out}}$, then
\begin{align*}
	g_0 z = \log\frac{E_{\textrm{out}}}{E} + \frac{E_{\textrm{out}}-E}{\Es},
\end{align*}
and by exponentiating, rearranging, and applying $W$, the Lambert $W$ function\footnote{see Appendix \ref{chap:lambertw}},
\begin{align*}
W \left(\frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 z}\right) &= W\left(\frac{E_{\textrm{out}}}{\Es} \textrm{e}^{E_{\textrm{out}}/\Es}\right) \\
&= \frac{E_{\textrm{out}}}{\Es},
\end{align*}
by \eqref{eq:lambertw}. This results in the closed form expression
\begin{align*}
E_{\textrm{out}}(z) = \Es W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 z} \right)
\end{align*}
with the desired property that $E_{\textrm{out}}(0)=E$. Furthermore, since $E \sim |A|^2$, the gain in terms of the amplitude is given by
\begin{align*}
G(A;E) &= \left(\frac{E_{\textrm{out}}(L_g)}{E}\right)^{1/2}A \\
&= \left( \frac{\Es}{E} W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 L_g} \right) \right)^{1/2} A,
\end{align*}
where $L_g$ is the length of the gain fibre.

\subsection{Fibre Nonlinearity}
The nonlinearity of the fibre arises from the parameter $\gamma$; in regions where this affect is dominant expression~\eqref{eq:nlse} becomes
\begin{align}
\label{eq:fibrediff}
	\pdiff{A}{z} - i \gamma |A|^2 A = 0.
\end{align}
This expression can be manipulated in a similar manner to the gain to show $\frac{\partial}{\partial z} |A|^2 = 0$, suggesting that $A(T,z) = A_0(T) e^{i \varphi(T,z)}$. Substituting this representation into~\eqref{eq:fibrediff} and setting $\varphi(T,0)=0$ gives $\varphi(T,z) = \gamma |A|^2 z$. For a fibre of length $L_f$ the effect of the nonlinearity is thus
\begin{align*}
F(A) &= A \textrm{e}^{i \gamma |A|^2 L_f},
\end{align*}
this is also frequently called the Kerr nonlinearity or Kerr effect after John Kerr who discovered it in 1875. As we shall see in Section \ref{chap:energy}, this is what is responsible for self-phase modulation and ultimately the degradation of the pulse \cite{desurvire, dunlop, martinez1984, tamura}.

@@

\subsection{Modulation}
In the average model, the amount of modulation is characterized by the parameter $M_s$ through the term $\frac{M_s}{2}T^2 A$. In this new model, the modulation is considered to be applied externally through its action on the spectrum and for simplicity the representation is taken as the Gaussian
\begin{align*}
M(A) &= \textrm{e}^{-T^2 / 2 T_M^2} A,
\end{align*}
where $T_M$ is a characteristic width of the modulation. \\

@@

\section{Solution to the Linear Model}
\label{sec:linear}
In the case of $b = 0$---which shall be referred to as the linear model---a solution can be found analytically. It is expected the solution will take the form of a Gaussian. There are a few reasons for this; the solution to the average model is a Gaussian, the equilibrium shape will be highly correlated to the shape of the modulation function, and since a Gaussian is a fixed point of the Fourier transform \cite{gradshteyn} dispersion will not alter the envelope. \\

@@

In equilibrium it must be that $A_0 = A_4 \textrm{e}^{-i \phi}$. More explicitly, this gives three conditions:
\begin{subequations}
\begin{align}
\label{eq:energycond}
1 &= \left( \frac{W(a E \textrm{e}^E)}{E} \right)^{1/2} h \zeta, \\
\label{eq:varcond}
\frac{1}{\sigma^2} &= \frac{1}{\widetilde{\sigma}^2} + 1, \\
\label{eq:chirpcond}
\frac{C}{\sigma^2} &= \frac{\widetilde{C}}{\widetilde{\sigma}^2}.
\end{align}
\end{subequations}

\subsection{Equilibrium Shape}
Each of these processes has a relatively straight forward effect, with the exception of dispersion. After dispersion, the out-going variance, $\widetilde{\sigma}^2$, can be found by computing the effect of dispersion given by \eqref{eq:disp}. This yields the relations
\begin{subequations}
\begin{align}
\label{eq:varout}
\widetilde{\sigma}^2 \sigma^2 &= \left( \sigma^2 + 2 C s^2 \right)^2 + 4s^4, \\
\label{eq:chirpout}
\widetilde{C} &= C + (1+C^2) \frac{2 s^2}{\sigma^2}, \\
\label{eq:phase}
\phi &= \frac{1}{2} \arctan \left( \frac{2s^2}{\sigma^2 + 2C s^2} \right), \\
\zeta &= \left( \frac{\sigma}{\widetilde{\sigma}} \right)^{1/2}
\end{align}
\end{subequations}
for the variance, chirp, phase shift, and spread, respectively, after the dispersive element, the full details can be found in Appendix \ref{chap:disp}. Now, the out-going variance and chirp can be eliminated from this system of equations using \eqref{eq:varcond} and \eqref{eq:chirpcond}:
\begin{align*}
\widetilde{\sigma}^2 &= \frac{\sigma^2}{1 - \sigma^2}, \\
\widetilde{C} &= C \frac{1}{1 - \sigma^2}.
\end{align*}
Combining this first expression with \eqref{eq:varout} and expanding, we have that
\begin{align*}
\frac{\sigma^4}{1 - \sigma^2} = \sigma^4 + 4 C^2 s^4 + 4 C s^2 \sigma^2 + 4s^4,
\end{align*}
or written as a polynomial in $\sigma$,
\begin{align*}
0 = \sigma^6 + 4 C s^2 \sigma^4 + \left( 4s^4 (C^2 + 1) - 4 C s^2 \right) \sigma^2 - 4s^4 (C^2 + 1).
\end{align*}
The chirp can now be eliminated using \eqref{eq:chirpout}, the $1+C^2$ can be reduced in order by noticing that
\begin{align*}
C \frac{1}{1 - \sigma^2} &= C + (1 + C^2) \frac{2s^2}{\sigma^2}, \\
1 + C^2 &= \frac{\sigma^4}{2s^2(1 - \sigma^2)} C.
\end{align*}
Furthermore, the chirp can be completely eliminated since
\begin{align}
\nonumber
C &= \frac{\sigma^4}{2s^2(1 - \sigma^2)} \pm \sqrt{\frac{\sigma^8}{16s^4(1 - \sigma^2)^2} - 1}, \\
\label{eq:equilchirp}
&= \frac{\sigma^4 \pm \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}}{4s^2(1 - \sigma^2)}.
\end{align}
After simplifying algebraically, we arrive at
\begin{align*}
0 &= \sigma^6 \pm \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}(2 - \sigma^2), \\
\frac{\sigma^6}{2 - \sigma^2} &= \mp \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}.
\end{align*}
As we shall see, $\sigma$ is strictly less than $1$ at equilibrium, and so, notice that the left hand side of this expression is strictly positive. Therefore, only the negative root of \eqref{eq:equilchirp} will yield a solution. Moreover, Martinez, Fork, and Gordon \cite{martinez1984} found two solutions analytically, but showed that one was unstable---consistent with experiments. After squaring each side of this expression we obtain
\begin{align*}
\sigma^{12} = \sigma^8 (2 - \sigma^2)^2 - 16s^4 (1 - \sigma^2)^2 (2 - \sigma^2)^2,
\end{align*}
which once fully simplified, yields the biquartic equation
\begin{align}
\label{eq:var}
\left( \sigma^2 \right)^4 + 4 s^4 \left( \sigma^2 \right)^3 - 20 s^4 \left( \sigma^2 \right)^2 + 32 s^4 \left( \sigma^2 \right) - 16 s^4 = 0.
\end{align}
Since this is a quartic in $\sigma^2$ this can be solved analytically; the (positive, real) solution is
\begin{align}
\label{eq:equilvar}
\sigma^2 &= \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} - s^4 - s^2 \sqrt{4 + s^4}.
\end{align}
The asymptotic expansions of the variance in the case of $s \rightarrow 0$, and $s \rightarrow \infty$ are derived in Appendix \ref{chap:asymp}. \\

>>>>

\chapter{A New Model}
\label{chap:linear}
In this chapter we shall derive our new model and solve it analytically in the linear case. To accomplish this we shall use the ideas of the previous functional models \cite{cutler, seigman, kuizenga1970a, kuizenga1970b, kuizenga1970, martinez1984, martinez1985, burgoyne2014}. To alleviate some of the pitfalls mentioned in the previous chapter, we shall include all five processes involved in the modification of the pulse within the laser cavity (gain, nonlinearity, loss, dispersion, and modulation). In addition to this, the functional operations associated with each component will be derived from \eqref{eq:nlse}, with the exception of the modulation in which we consider the exact functional form to be determined by the laser operator. \\

\section{Components}
We shall begin our analysis with the derivation of the functional operators for the five components. \\

\subsection{Gain}
\label{chap:gain}
Within the Er-doped gain fibre, the gain term is dominant, and equation~\eqref{eq:nlse} reduces to
\begin{align}
\label{eq:gainde}
\pdiff{A}{z} &= \frac{1}{2} g(A),
\end{align}
where $g(A)$ takes the form \cite{bohun, burgoyne2014, hausbook, haus1975, haus1991, haus1992, haus2000, kartner, peng, shtyrina, silfvast, usechak, yarutkina}
\begin{align}
\label{eq:energy}
	g(A) &= \frac{g_0}{1 + E / \Es} A,& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate.
Multiplying~\eqref{eq:gainde} by $\bar{A}$, the complex conjugate of $A$, yields
\begin{align*}
	\bar{A} \pdiff{A}{z} = \frac{1}{2}\frac{g_0 |A|^2}{1 + E / \Es},
\end{align*}
adding this to its complex conjugate and integrating over $T$ gives
%\begin{align*}
%	\diff{|A|^2}{z} &= \frac{g_0 |A|^2}{1 + E / \Es}.
%\end{align*}
%After integrating this becomes
\begin{align}
\label{diffez}
	\diff{E}{z} &= \frac{g_0 E}{1 + E / \Es}.
\end{align}
For $E \ll \Es$ the energy will grow exponentially, whereas for $E \gg \Es$ the gain has saturated and so the growth is linear. To obtain a closed form solution, \eqref{diffez} is integrated over a gain fibre of length $z$ and assuming the energy increases from $E$ to $E_{\textrm{out}}$, then
\begin{align*}
	g_0 z = \log\frac{E_{\textrm{out}}}{E} + \frac{E_{\textrm{out}}-E}{\Es},
\end{align*}
and by exponentiating, rearranging, and applying $W$, the Lambert $W$ function\footnote{see Appendix \ref{chap:lambertw}},
\begin{align*}
W \left(\frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 z}\right) &= W\left(\frac{E_{\textrm{out}}}{\Es} \textrm{e}^{E_{\textrm{out}}/\Es}\right) \\
&= \frac{E_{\textrm{out}}}{\Es},
\end{align*}
by \eqref{eq:lambertw}. This results in the closed form expression
\begin{align*}
E_{\textrm{out}}(z) = \Es W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 z} \right)
\end{align*}
with the desired property that $E_{\textrm{out}}(0)=E$. Furthermore, since $E \sim |A|^2$, the gain in terms of the amplitude is given by
\begin{align*}
G(A;E) &= \left(\frac{E_{\textrm{out}}(L_g)}{E}\right)^{1/2}A \\
&= \left( \frac{\Es}{E} W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 L_g} \right) \right)^{1/2} A,
\end{align*}
where $L_g$ is the length of the gain fibre.

\subsection{Fibre Nonlinearity}
The nonlinearity of the fibre arises from the parameter $\gamma$; in regions where this affect is dominant expression~\eqref{eq:nlse} becomes
\begin{align}
\label{eq:fibrediff}
	\pdiff{A}{z} - i \gamma |A|^2 A = 0.
\end{align}
This expression can be manipulated in a similar manner to the gain to show $\frac{\partial}{\partial z} |A|^2 = 0$, suggesting that $A(T,z) = A_0(T) e^{i \varphi(T,z)}$. Substituting this representation into~\eqref{eq:fibrediff} and setting $\varphi(T,0)=0$ gives $\varphi(T,z) = \gamma |A|^2 z$. For a fibre of length $L_f$ the effect of the nonlinearity is thus
\begin{align*}
F(A) &= A \textrm{e}^{i \gamma |A|^2 L_f},
\end{align*}
this is also frequently called the Kerr nonlinearity or Kerr effect after John Kerr who discovered the effect in 1875. As we shall see in Section \ref{chap:energy}, this is what is responsible for self-phase modulation and ultimately the degradation of the pulse \cite{desurvire, dunlop, martinez1984, tamura}.

@@

\subsection{Modulation}
In the master equation of mode-locking, the amount of modulation is characterized by the parameter $M_s$ through the term $\frac{M_s}{2}T^2 A$. In this new model, the modulation is considered to be applied externally through its action on the spectrum and for simplicity the representation is taken as the Gaussian
\begin{align*}
M(A) &= \textrm{e}^{-T^2 / 2 T_M^2} A,
\end{align*}
where $T_M$ is a characteristic width of the modulation. \\

@@

\section{Solution to the Linear Model}
\label{sec:linear}
In the case of $b = 0$---which shall be referred to as the linear model---a solution can be found analytically. It is expected the solution will take the form of a Gaussian. There are a few reasons for this; the solution to the previous models were Gaussian \cite{cutler, seigman, kuizenga1970a, martinez1984, martinez1985}, the equilibrium shape will be highly correlated to the shape of the modulation function, and since a Gaussian is a fixed point of the Fourier transform \cite{gradshteyn} dispersion will not alter the envelope. \\

@@

In equilibrium it must be that $A_0 = A_4 \textrm{e}^{-i \phi}$. More explicitly, this gives three conditions:
\begin{subequations}
\begin{align}
\label{eq:energycond}
1 &= \left( \frac{W(a E \textrm{e}^E)}{E} \right)^{1/2} h \zeta, \\
\label{eq:varcond}
\frac{1}{\sigma^2} &= \frac{1}{\widetilde{\sigma}^2} + 1, \\
\label{eq:chirpcond}
\frac{C}{\sigma^2} &= \frac{\widetilde{C}}{\widetilde{\sigma}^2}.
\end{align}
\label{eq:linsys}
\end{subequations}

\subsection{Dispersion}
Each of these processes has a relatively straight forward effect, with the exception of dispersion.
\input{./Appendices/Dispersion}

\subsection{Equilibrium Shape}
Now that the effect of dispersion is know analytically, the system \eqref{eq:linsys} can be solved. The out-going variance and chirp can be eliminated from this system of equations using \eqref{eq:varcond} and \eqref{eq:chirpcond}:
\begin{align*}
\widetilde{\sigma}^2 &= \frac{\sigma^2}{1 - \sigma^2}, \\
\widetilde{C} &= C \frac{1}{1 - \sigma^2}.
\end{align*}
Combining this first expression with \eqref{eq:varout} and expanding, we have that
\begin{align*}
\frac{\sigma^4}{1 - \sigma^2} = \sigma^4 + 4 C^2 s^4 + 4 C s^2 \sigma^2 + 4s^4,
\end{align*}
or written as a polynomial in $\sigma$,
\begin{align*}
0 = \sigma^6 + 4 C s^2 \sigma^4 + \left( 4s^4 (C^2 + 1) - 4 C s^2 \right) \sigma^2 - 4s^4 (C^2 + 1).
\end{align*}
The chirp can now be eliminated using \eqref{eq:chirpout}, the $1+C^2$ can be reduced in order by noticing that
\begin{align*}
C \frac{1}{1 - \sigma^2} &= C + (1 + C^2) \frac{2s^2}{\sigma^2}, \\
1 + C^2 &= \frac{\sigma^4}{2s^2(1 - \sigma^2)} C.
\end{align*}
Furthermore, the chirp can be completely eliminated since
\begin{align}
\nonumber
C &= \frac{\sigma^4}{2s^2(1 - \sigma^2)} \pm \sqrt{\frac{\sigma^8}{16s^4(1 - \sigma^2)^2} - 1}, \\
\label{eq:equilchirp}
&= \frac{\sigma^4 \pm \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}}{4s^2(1 - \sigma^2)}.
\end{align}
After simplifying algebraically, we arrive at
\begin{align*}
0 &= \sigma^6 \pm \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}(2 - \sigma^2), \\
\frac{\sigma^6}{2 - \sigma^2} &= \mp \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}.
\end{align*}
As we shall see, $\sigma$ is strictly less than $1$ at equilibrium, and so, notice that the left hand side of this expression is strictly positive. Therefore, only the negative root of \eqref{eq:equilchirp} will yield a solution. Moreover, Martinez, Fork, and Gordon \cite{martinez1984} found two solutions analytically, but showed that one was unstable---consistent with experiments. After squaring each side of this expression we obtain
\begin{align*}
\sigma^{12} = \sigma^8 (2 - \sigma^2)^2 - 16s^4 (1 - \sigma^2)^2 (2 - \sigma^2)^2,
\end{align*}
which once fully simplified, yields the biquartic equation
\begin{align}
\label{eq:var}
\left( \sigma^2 \right)^4 + 4 s^4 \left( \sigma^2 \right)^3 - 20 s^4 \left( \sigma^2 \right)^2 + 32 s^4 \left( \sigma^2 \right) - 16 s^4 = 0.
\end{align}
Since this is a quartic in $\sigma^2$ this can be solved analytically; the (positive, real) solution is
\begin{align}
\label{eq:equilvar}
\sigma^2 &= \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} - s^4 - s^2 \sqrt{4 + s^4}.
\end{align}

\subsection{Asymptotic Expansion of the Variance}
While \eqref{eq:equilvar} is useful, so too is the asymptotic behaviour.
\input{./Appendices/Asymptotics}
